### 2016-07-27 syntax.tex
updating syntax of data structures, and adding set-multiplicity data structures

\section{Secure Data Structures}
\def\pub{{\sf pub}}
\def\aux{{\sf aux}}

@@

@@

\heading{Examples.}
It is not hard to see Bloom filters are one possible instantiation a set-multiplicity data structure.
Fix an $n \geq 0$, let $\mathcal{S}=[\univ]^n$, and let $\Sigma=\Gamma\bits$.  
Let $\calQ$ be a set of predicates $q_x$, one for each $x \in \univ$,
such that $q_x(S)=1$ iff $x \in S$.  Define algorithm~$\Rep$, with functions $k=k(|S|), m=m(|S|)$ and a description of a hash function family~$\mathcal{H}=\{h \colon \univ \to \{1,2,\ldots,m\}\}$ hardwired into it, to sample hash functions $h_1, \ldots, h_k$ from~$\mathcal{H}$ and then computes an $m$-bit array in the standard way. It sets $\aux = (h_1, \ldots, h_k)$ and $M$ to be the bit array, and returns these.  The algorithm~$\Qry$ is defined so that, on input $(M,\aux,q_x)$, it returns 1 iff $M[h_1(x)]=1 \wedge M[h_2(x)]=1 \wedge \cdots \wedge M[h_k(x)]=1$ holds.

We note that, while our measure of error~$\epsilon$ does not distinguish among ``types'' of errors, it also does not preclude instantiations (like the classical Bloom filter) that admit only one type of error, e.g. false-positives.

A Bloom filter with secret hash functions is captured by appending to~$M$ appropriate encodings of $h_1,h_2,\ldots,h_k$, and setting $\aux=\emptystring$.  Related constructions that use a secret key~$K$, but not for hashing are captured in analogous ways.

The so-called ``garbled Bloom filter'' due to Dong, Chen and Wen~\cite{xxx}, which stores xor-shares of bitstring~$x \in S$ at positions $h_1(x),h_2(x),\ldots,h_k(x)$ in an array, falls within our syntax by letting~$\Sigma=\bits^L$ for a specified integer~$L>0$.  Similarly the cuckoo-hashing construction from NY, which (loosely speaking) stores a hash $g(x)$ at positions determined by $h_1(x),h_2(x),\ldots,h_k(x)$. \tsnote{Check this, I forget the exact details now.}

Expanding beyond queries that are predicates, consider $\Sigma=\bits^L$ and a scheme that stores $L$-bit counters at each position in an array.  In particular, for each~$x \in S$ the $\Rep$~algorithm increments the counters stored at $h_1(x),h_2(x),\ldots,h_k(x)$, returns the final array of counters as~$M$.  This representation naturally supports a set of queries $\calQ$ where $q_x$ has the semantics ``what is the multiplicity of~$x$ in~$S$.''   Such a construction gives a static version of a counting Bloom filter~\cite{xxx}.  (We will capture dynamic versions, as well as count-min sketches, scalable Bloom filters, etc., in a later section.)

>>>>

\section{(Set-multiplicity) Data Structures}
\def\pub{{\sf pub}}
\def\aux{{\sf aux}}

@@

\todo{Add discussion of just how broad the syntax is, in general.}

@@

\heading{Bloom Filters and Related Data Structures. }
It is not hard to see Bloom filters are one possible instantiation a set-multiplicity data structure.
Fix an $n \geq 0$, let $\mathcal{S}=[\univ]^n$, and let $\Sigma=\Gamma=\bits$.  
Define the queries so that $q_x(S)=1$ if $x \in S$, and 0 otherwise.  Define algorithm~$\Rep$, with functions $k=k(|S|), m=m(|S|)$ and a description of a hash function family~$\mathcal{H}=\{h \colon \univ \to \{1,2,\ldots,m\}\}$ hardwired into it, to sample hash functions $h_1, \ldots, h_k$ from~$\mathcal{H}$ and then compute an $m$-bit array in the standard way.  It sets $\aux = (h_1, \ldots, h_k)$ and $M$ to be the bit array, and returns these.  The algorithm~$\Qry$ is defined so that, on input $(M,\aux,q_x)$, it returns the minimum of the values (either 0 or 1) held at $M[h_1(x)],\ldots,M[h_k(x)]$.  We note that, while our error-probability and error-rate do not distinguish among ``types'' of errors, they also do not preclude instantiations (like the classical Bloom filter) that admit only one type of error, e.g. false-positives.

As a simple extension, consider $\Sigma=\mathbb{N}$ and a scheme that stores counters at each position in an array.  In particular, for each~$x \in S$ the $\Rep$~algorithm increments the counters stored at positions $h_1(x), h_2(x), \ldots, h_k(x)$, returns the final array of counters as~$M$.  The $\Qry$-algorithm remains the same. Such a construction gives a static version of a counting Bloom filter~\cite{xxx}.  We will capture dynamic versions, as well as count-min sketches, scalable Bloom filters, etc., in a later section.

A Bloom filter with secret hash functions is captured by appending to~$M$ appropriate encodings of $h_1,h_2,\ldots,h_k$, and setting $\aux=\emptystring$.  Related constructions \tsnote{Name them?} that use a secret key~$K$, but not for hashing, are captured by appending~$K$ to~$M$ and setting $\aux=((h_1, \ldots, h_k)$.

The so-called ``garbled Bloom filter'' due to Dong, Chen and Wen~\cite{xxx}, which stores xor-shares of bitstring~$x \in S$ at positions $h_1(x),h_2(x),\ldots,h_k(x)$ in an array, falls within our syntax by letting~$\Sigma=\bits^L$ for a specified integer~$L>0$.  Similarly the cuckoo-hashing construction from NY, which (loosely speaking) stores a hash $g(x)$ at positions determined by $h_1(x),h_2(x),\ldots,h_k(x)$. \tsnote{Check the correctness of this claim.}
